### 2021-09-06 submission/manuscript.tex
add sample size restriction for adaptive method

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{graphicx}

\title{A model including an interaction of treatment with baseline risk
was the optimal risk-based approach to predicting absolute treatment
benefit}

\author{
    Alexandros Rekkas
   \\
    Department of Medical Informatics \\
    Erasmus Medical Center \\
  Rotterdam, The Netherlands \\
  \texttt{} \\
   \And
    Peter R. Rijnbeek
   \\
    Department of Medical Informatics \\
    Erasmus Medical Center \\
  Rotterdam, The Netherlands \\
  \texttt{} \\
   \And
    Ewout W. Steyerberg
   \\
    Department of Biomedical Data Sciences \\
    Leiden University Medical Center \\
  Leiden, The Netherlands \\
  \texttt{} \\
   \And
    David van Klaveren
   \\
    Department of Public Health \\
    Erasmus Medical Center \\
  Rotterdam, The Netherlands \\
  \texttt{} \\
  }

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\usepackage{amsmath}
\doublespacing
\usepackage[left]{lineno}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{booktabs}
\newcommand\given[1][]{\:#1\vert\:}
\date{}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\maketitle

\def\tightlist{}

\begin{abstract}
\textbf{Objective:} To compare different risk-based methods predicting
individualized treatment effects in RCT simulations. \textbf{Study
Design and Setting:} We simulated data using diverse assumptions for a
baseline prognostic index of risk (PI) and the shape of its
intereraction with treatment (none, linear or quadratic). In each sample
we predicted absolute benefit using: models with the PI and a constant
relative treatment effect; models including an interaction of treatment
with the PI; stratification in quarters of the PI; nonlinear
transformations of the PI (restricted cubic splines with 3, 4 and 5
knots); an adaptive approach using Akaike's Information Criterion. We
evaluated predictive performance using root mean squared error and
measures of discrimination and calibration for benefit. Starting from a
base case scenario (sample size 4,250, treatment odds ratio 0.8, AUC of
the PI 0.75), we varied the sample size, the treatment effect strength,
and the PI's discriminative ability. \textbf{Results:} Models including
a PI by treatment interaction performed better under most simulation
settings. Flexible models required larger sample sizes and higher AUC of
the PI to outperform the linear interaction model. The adaptive approach
performed similarly to the best-performing method in most scenarios.
\textbf{Conclusion:} Under most circumstances, a model with a linear
interaction of the PI with treatment is the optimal risk-based approach
to predicting absolute treatment benefit.
\end{abstract}

\keywords{
    treatment effect heterogeneity
   \and
    absolute benefit
   \and
    prediction models
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Predictive approaches for assessing heterogeneity of treatment effects
(HTE) aim at the development of models predicting either individualized
effects or which of two (or more) treatments is better for an individual
{[}1{]}. In prior work, we divided such methods in three broader
categories based on the reference class used for defining patient
similarity when making individualized predictions or recommendations
{[}2{]}. Risk-modeling approaches use prediction of baseline risk as the
reference; treatment effect modeling approaches also model
treatment-covariate interactions, in addition to risk factors; optimal
treatment regime approaches focus on developing treatment assignment
rules and therefore rely heavily on modeling treatment effect modifiers.

Risk-modeling approaches to predictive HTE analyses provide a viable
option in the absence of well-established treatment effect modifiers
{[}3,4{]}. In simulations, modeling of effect modifiers, i.e.
treatment-covariate interactions, often led to miscalibrated predictions
of benefit, while risk-based methods proved quite robust {[}5{]}. Most
often, risk-modeling approaches are carried out in two steps: first a
risk prediction model is developed externally or internally on the
entire RCT population, ``blinded'' to treatment; then the RCT population
is stratified using this prediction model to evaluate risk-based
treatment effect variation {[}6{]}. However, even though estimates at
the risk subgroup level may be accurate, these estimates do not apply to
individual patients, especially for patients with predicted risk at the
boundaries of the risk intervals. Hence, the risk-stratified approach is
useful for exploring and presenting HTE, but is not useful for
supporting treatment decisions for individual patients.

To individualize treatment effects, the recent PATH statement suggested
various risk-based models including a prognostic index of baseline risk
(PI) and treatment assignment {[}3,4{]}. We aimed to summarize and
compare different risk-based models for predicting individualized
treatment effects. We simulated RCT settings to compare the performance
of these models under different assumptions of the relationship between
baseline risk and treatment. We illustrated the different models by a
case study of predicting individualized effects of tissue plasminogen
activator (tPA) versus streptokinase treatment in patients with an acute
myocardial infarction (MI).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{simulation-scenarios}{%
\subsection{Simulation scenarios}\label{simulation-scenarios}}

For each patient we generated 8 baseline covariates here
\(x_1,\dots,x_4\sim N(0, 1)\) and \(x_5,\dots,x_8\sim B(1, 0.2)\).
Treatment was allocated using a 50:50 split. Outcomes for patients in
the control arm were generated from a logistic regression model
including all baseline covariates. In the base case scenario coefficient
values were such, that the AUC of the logistic regression model was
\(0.75\) and the event rate in the control arm was \(20\%\).

While binary outcomes in the control arm were generated from Bernoulli
variables with true probabilities
\(P(y=1|X) = \text{expit}(PI)=\frac{e^{PI}}{1+e^{PI}}\), the outcomes in
the treatment arm were based on true probabilities
\(\text{expit}(lp_1)\), with
\[lp_1 = \gamma_2(PI-c)^2 + \gamma_1(PI-c) + \gamma_0\]

The coefficients \(\gamma_0,\gamma_1\) and \(\gamma_2\) along with the
centering constant \(c\) were set for each simulation scenario to mimic
a wide variety scenarios, ranging from true constant relative treatment
effect (\(\gamma_1=\gamma_2=0\)) to moderate and strong linear
(\(\gamma_2=0\)) and quadratic deviations. We also considered scenarios
with treatment-covariate interactions. These scenarios included 4 weak
interactions (\(\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.82\)), 4 strong
interactions (\(\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.61\)), and 2
weak and 2 strong interactions.

The sample size of the base case scenario was set to 4,250 (\(80\%\)
power for the detection of a marginal OR of 0.8). We evaluated the
effect of smaller or larger sample sizes of 1,064 and 17,000,
respectively. We also evaluated the effect of worse or better
discriminative ability for risk, adjusting the baseline covariate
coefficients, such that the AUC of the regression model in the control
arm was 0.65 and 0.85 respectively.

Combining all these settings resulted in a simulation study of 66
scenarios (exact settings in the supplementary material).

\hypertarget{individualized-risk-based-benefit-predictions}{%
\subsection{Individualized risk-based benefit
predictions}\label{individualized-risk-based-benefit-predictions}}

All methods assume that a risk prediction model is available to assign
risk predictions to individual patients. For the simulations we
developed a prediction model internally, using logistic regression
including main effects for all baseline covariates and treatment
assignment. Risk predictions for individual patients were based on
treatment assignment to the control arm, that is setting treatment
assignment to 0.

A \emph{stratified HTE method} has been suggested as an alternative to
traditional subgroup analyses. Patients are stratified into
equally-sized risk strata---in this case based on risk quartiles.
Absolute treatment effects within risk strata are estimated by the
difference in event rate between patients in the control arm and
patients in the treated arm. We considered this approach as a reference,
expecting it to perform worse than the other candidates, as its
objective is not to individualize benefit prediction.

Second, we considered a model which assumes \emph{constant relative
treatment effect} (constant OR). Hence, absolute benefit is predicted
from \(\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))\).

Third, we considered a logistic regression model including treatment,
the prognostic index, and their linear interaction. Absolute benefit is
then estimated from
\(\hat{\tau}(\bm{x})=\text{expit}(\beta_0+\beta_{PI}PI) - \text{expit}(\beta_0+\beta_{t_x}+(\beta_{PI}+\beta_*)PI)\).
We will refer to this method as the \emph{linear interaction} approach.

Fourth, we used \emph{restricted cubic splines} (RCS) to relax the
linearity assumption on the effect of the linear predictor {[}7{]}. We
considered splines with 3, 4 and 5 knots to compare models with
different levels of flexibility.

Finally, we considered an adaptive approach using Akaike's Information
Criterion (AIC) for model selection. The candidate models were: a
constant treatment effect model, a model with a linear interaction with
treatment and RCS models with 3, 4 and 5 knots.

\hypertarget{evaluation-metrics}{%
\subsection{Evaluation metrics}\label{evaluation-metrics}}

We evaluated the predictive accuracy of the considered methods by the
root mean squared error (RMSE):

\[\text{RMSE}=\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2\]
We compared the discriminative ability of the methods under study using
c-for-benefit {[}8{]}. The c-for-benefit represents the probability that
from two randomly chosen matched patient pairs with unequal observed
benefit, the pair with greater observed benefit also has a higher
predicted benefit. To be able to calculate observed benefit, patients in
each treatment arm are ranked based on their predicted benefit and then
matched 1:1 across treatment arms. \emph{Observed} treatment benefit is
defined as the difference of observed outcomes between the untreated and
the treated patient of each matched patient pair. \emph{Predicted}
benefit is defined as the average of predicted benefit within each
matched patient pair.

We evaluated calibration in a similar manner, using the integrated
calibration index (ICI) for benefit {[}9{]}. The observed benefits are
regressed on the predicted benefits using a locally weighted scatterplot
smoother (loess). The ICI-for-benefit is the average absolute difference
between predicted and smooth observed benefit. Values closer to \(0\)
represent better calibration.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{simulations}{%
\subsection{Simulations}\label{simulations}}

The model including a constant relative treatment effect had the lowest
median RMSE in scenarios with a true constant relative treatment effect
(OR = 0.8, N = 4,250 and AUC = 0.75) or with moderate relative
deviations (Figure \ref{fig:rmsebase}; Panel A). However, when we
considered strong linear and quadratic deviations from a constant
relative treatment effect the linear interaction model performed best
(Figure \ref{fig:rmsebase}; Panels B and C). Only in the case of strong
quadratic deviations models including RCS (3 knots) performed equally
well to the linear interaction method. Increasing the number of knots in
RCS resulted in higher RMSE across all scenarios. The adaptive approach
performed very similarly to the best performing model in each scenario.

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_base} \caption{RMSE of the considered methods across 500 replications calculated in a simulated superpopulation of size 500,000. The scenario with true constant relative treatment effect had a true prediction AUC of 0.75 and sample size of 4250. Results are presented under moderate linear, strong linear, and strong quadratic deviations from constant relative treatment effects.}\label{fig:rmsebase}
\end{figure}

When we increased the sample size (N = 17,000), the model including a
constant relative treatment effect had the lowest RMSE under the
assumption of true constant relative treatment effects (Figure
\ref{fig:rmsesamplesize}; Panel A). However, when introducing moderate
and strong linear deviations the linear interaction model outperformed
the constant treatment effect model (Figure \ref{fig:rmsesamplesize};
Panels B and C). Furthermore, the linear interaction model was
outperformed by the more flexible RCS models (3 knots) in the case of
strong quadratic deviations. Again, the increased flexibility of RCS
smoothing with higher number of knots resulted in overfitting and worse
predictive accuracy (Figure \ref{fig:rmsesamplesize}; Panel D).

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_sample_size} \caption{RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. Sample size 17,000 rather than 4250 in Figure \ref{fig:rmsebase}}\label{fig:rmsesamplesize}
\end{figure}

When we increased the AUC of the true prediction model to 0.85, models
including RCS smoothing had the lowest RMSE in the presence of strong
quadratic deviations from the base case of true constant relative
treatment effects (Figure \ref{fig:rmseauc}; Panel D). However, with
milder deviations, the linear interaction model had the lowest RMSE with
the RCS smoothing methods (3 knots) being a close second (Figure
\ref{fig:rmseauc}; Panels B and C). Increasing the number of knots of
RCS smoothing resulted in increased RMSE, which was less pronounced in
the case of strong quadratic deviations.

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_auc} \caption{RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.85 and sample size of 4,250.}\label{fig:rmseauc}
\end{figure}

The constant effects model, the linear interaction model and models with
RCS smoothing (3 knots) had the highest median c-for-benefit in the base
case scenario and the scenarios where linear and quadratic deviations
were considered. The constant treatment effect model and the linear
interaction model tended to present much lower variability compared to
all other approaches (Figure \ref{fig:discrimination}). With increasing
number of RCS knots, we observed decreasing median values and increasing
variability of the c-for-benefit in all scenarios.

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/discrimination_base} \caption{Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.75 and sample size of 4,250.}\label{fig:discrimination}
\end{figure}

As for calibration, the linear interaction model had the lowest median
ICI-for-benefit in the majority of scenarios except for the scenarios
where no or moderate linear deviations from the base case were
considered. In that case constant treatment effect models demonstrated
optimal calibration, very comparable to the linear interaction model's
performance, nonetheless (Figure \ref{fig:calibration}). With strong
linear or quadratic deviations, the constant treatment effect model was
poorly calibrated (Figure \ref{fig:calibration}; Panels C and D).

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/calibration_base} \caption{Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.75 and sample size of 4,250.}\label{fig:calibration}
\end{figure}

\hypertarget{case-study}{%
\subsection{Case study}\label{case-study}}

We demonstrate the different methods for individualizing treatment
benefits using data from 30,510 patients with an acute myocardial
infarction (MI) included in the GUSTO-I trial. 10,348 patients were
randomized to tissue plasminogen activator (tPA) treatment and 20,162
were randomized to streptokinase. The outcome of interest was 30-day
mortality, recorded for all patients.

In line with previous analyses {[}10,11{]}, we fitted a logistic
regression model with 6 baseline covariates, i.e.~age, Killip class,
systolic blood pressure, heart rate, an indicator of previous MI, and
the location of MI, to predict 30-day mortality risk. A constant effect
of treatment was included in the model. When deriving risk predictions
for individuals we set the treatment indicator to 0. More information on
model development can be found in the supplement.

We used the risk linear predictor to fit the proposed methods under
study for individualizing absolute benefit predictions. All methods
predicted increasing benefits for patients with higher baseline risk
predictions, but the fitted patterns were clearly different. The
adaptive approach selected the model with RCS smoothing with 4 knots.
However, for very low baseline risk this model predicted decreasing
benefit with increasing risk may be somewhat too flexible. The more
robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice. The linear interaction model
had somewhat lower AIC compared to the model with RCS smoothing (3
knots), slightly better cross-validated discrimination (c-for-benefit
0.526 vs 0.525) and quite similar cross-validated calibration (ICI-for
benefit 0.0115 vs 0.0117).

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/gusto} \caption{Individualized absolute benefit predictions based on baseline risk when using a constantn treatment effect approach, a linear interaction approach and RCS smoothing using 3,4 and 5 knots. Risk stratified estimates of absolute benefit are presented within quartiles of baseline risk as reference.}\label{fig:gusto}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The linear interaction model proved to be superior to alternative
approaches when predicting risk-based treatment benefit under a wide
range of scenarios. It generally had lower mean squared error, lower
e-for-benefit and higher c-for-benefit with lower variability across
simulation replications. Models including restricted cubic splines with
3 knots only outperformed the linear interaction model in the presence
of strong quadratic deviations from a constant relative treatment
effect.

Models including restricted cubic splines with 4 or 5 knots proved to be
too flexible, as indicated by higher RMSE, increased variability of
discrimination for benefit and worse calibration of benefit predictions.
Even with larger sample sizes and strong quadratic deviations from the
base case scenario of constant relative treatment effects, these more
flexible restricted cubic splines did not outperform the simpler RCS
with 3 knots. These approaches may only be helpful if we expect more
extreme patterns of heterogeneous treatment effects compared to the
quadratic deviations considered here.

The constant treatment effect model, despite having adequate performance
in the presence of weak treatment effect heterogeneity on the relative
scale, quickly broke down with stronger deviations from constant
relative treatment effects. In these cases, the stratified approach
generally had lower error rates compared to the constant treatment
effect model. Stepwise treatment benefit estimates are very useful for
demonstrating treatment effect heterogeneity--because estimating
treatment effect requires groups of patients rather than individual
patients--but are not helpful for making individualized absolute benefit
predictions.

Increasing the discriminative ability of the risk model--by increasing
the predictor coefficients of the true risk model--reduced RMSE for all
methods. This increase in discriminative ability translates in higher
variability of predicted risks, which, in turn, allows the considered
methods to better capture absolute treatment benefits. As a consequence,
the increase in discriminative ability of the risk model also led to
higher values of c-for-benefit. Even though risk model performance is
very important for the ability of risk-based methods to predict
treatment benefit, prediction model development was outside the scope of
this work and has already been studied extensively {[}5,12,13{]}.

The adaptive approach had adequate performance, following closely the
performance of the ``true'' model in most scenarios. However, with
smaller sample sizes it tended to ``miss'' the treatment-risk
interactions and selected simpler models. This resulted in increased
RMSE variability in these scenarios, especially in the case of true
strong linear deviations from the base case scenario. This tendency was
mitigated when the discriminative ability of the prediction model was
higher. Therefore, in the case of smaller sample sizes and/or poorly
discriminating prediction models, the simpler linear interaction model
is a safer choice for predicting absolute benefits.

Risk-based approaches to predictive HTE estimate treatment benefit as a
function of baseline risk. A limitation of our study is that we assumed
treatment benefit to be a function of baseline risk in the majority of
the simulation scenarios. Nevertheless, our main conclusions did not
change when we generated individual treatment-covariate interactions
(supplementary table/figure x). Future simulation studies could explore
the effect of more extensive deviations from risk-based treatment
effects.

Recent years have seen an increased interest in predictive HTE
approaches focusing on individualized benefit predictions. In our
simulations we only focused on risk-based methods, using baseline risk
as a reference in a two-stage approach to individualizing benefit
predictions. However, there is a plethora of different methods, ranging
from treatment effect modeling to tree-based approaches available in
more recent literature {[}14--16{]}. Simulations are also needed to
assess relative performance and define the settings where these break
down or outperform each other.

In conclusion, when comparing different risk-based approaches to
predicting individualized treatment benefit, a model including a linear
treatment interaction with the prognostic index performed best in a wide
range of scenarios. More flexible models with restricted cubic splines
required larger sample sizes and higher AUC of the prognostic index to
outperform the linear interaction model. An adaptive approach, selecting
the model with the optimal AIC, had comparable performance to the best
performing approach in most of the scenarios.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}

\noindent

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-Varadhan2013}{}%
{[}1{]} Varadhan R, Segal JB, Boyd CM, Wu AW, Weiss CO. A framework for
the analysis of heterogeneity of treatment effect in~patient-centered
outcomes research. Journal of Clinical Epidemiology 2013;66:818--25.
\url{https://doi.org/10.1016/j.jclinepi.2013.02.009}.

\leavevmode\hypertarget{ref-Rekkas2020}{}%
{[}2{]} Rekkas A, Paulus JK, Raman G, Wong JB, Steyerberg EW, Rijnbeek
PR, et al. Predictive approaches to heterogeneous treatment effects: A
scoping review. BMC Medical Research Methodology 2020;20.
\url{https://doi.org/10.1186/s12874-020-01145-1}.

\leavevmode\hypertarget{ref-Kent2019}{}%
{[}3{]} Kent DM, Paulus JK, Klaveren D van, D'Agostino R, Goodman S,
Hayward R, et al. The predictive approaches to treatment effect
heterogeneity (PATH) statement. Annals of Internal Medicine 2019;172:35.
\url{https://doi.org/10.7326/m18-3667}.

\leavevmode\hypertarget{ref-PathEnE}{}%
{[}4{]} Kent DM, Klaveren D van, Paulus JK, D'Agostino R, Goodman S,
Hayward R, et al. The predictive approaches to treatment effect
heterogeneity (PATH) statement: Explanation and elaboration. Annals of
Internal Medicine 2019;172:W1. \url{https://doi.org/10.7326/m18-3668}.

\leavevmode\hypertarget{ref-vanKlaveren2019}{}%
{[}5{]} Klaveren D van, Balan TA, Steyerberg EW, Kent DM. Models with
interactions overestimated heterogeneity of treatment effects and were
prone to treatment mistargeting. Journal of Clinical Epidemiology
2019;114:72--83. \url{https://doi.org/10.1016/j.jclinepi.2019.05.029}.

\leavevmode\hypertarget{ref-Kent2010}{}%
{[}6{]} Kent DM, Rothwell PM, Ioannidis JP, Altman DG, Hayward RA.
Assessing and reporting heterogeneity in treatment effects in clinical
trials: A proposal. Trials 2010;11.
\url{https://doi.org/10.1186/1745-6215-11-85}.

\leavevmode\hypertarget{ref-Harrell1988}{}%
{[}7{]} Harrell FE, Lee KL, Pollock BG. Regression models in clinical
studies: Determining relationships between predictors and response. JNCI
Journal of the National Cancer Institute 1988;80:1198--202.
\url{https://doi.org/10.1093/jnci/80.15.1198}.

\leavevmode\hypertarget{ref-vanKlaveren2018}{}%
{[}8{]} Klaveren D van, Steyerberg EW, Serruys PW, Kent DM. The proposed
``concordance-statistic for benefit'' provided a useful metric when
modeling heterogeneous treatment effects. Journal of Clinical
Epidemiology 2018;94:59--68.
\url{https://doi.org/10.1016/j.jclinepi.2017.10.021}.

\leavevmode\hypertarget{ref-Austin2019}{}%
{[}9{]} Austin PC, Steyerberg EW. The integrated calibration index (ICI)
and related metrics for quantifying the calibration of logistic
regression models. Statistics in Medicine 2019;38:4051--65.
\url{https://doi.org/10.1002/sim.8281}.

\leavevmode\hypertarget{ref-Califf1997}{}%
{[}10{]} Califf RM, Woodlief LH, Harrell FE, Lee KL, White HD, Guerci A,
et al. Selection of thrombolytic therapy for individual patients:
Development of a clinical model. American Heart Journal 1997;133:630--9.
\url{https://doi.org/10.1016/s0002-8703(97)70164-9}.

\leavevmode\hypertarget{ref-Steyerberg2000}{}%
{[}11{]} Steyerberg EW, Bossuyt PMM, Lee KL. Clinical trials in acute
myocardial infarction: Should we adjust for baseline characteristics?
American Heart Journal 2000;139:745--51.
\url{https://doi.org/10.1016/s0002-8703(00)90001-2}.

\leavevmode\hypertarget{ref-Burke2014}{}%
{[}12{]} Burke JF, Hayward RA, Nelson JP, Kent DM. Using internally
developed risk models to assess heterogeneity in treatment effects in
clinical trials. Circulation: Cardiovascular Quality and Outcomes
2014;7:163--9. \url{https://doi.org/10.1161/circoutcomes.113.000497}.

\leavevmode\hypertarget{ref-Abadie2018}{}%
{[}13{]} Abadie A, Chingos MM, West MR. Endogenous stratification in
randomized experiments. The Review of Economics and Statistics
2018;100:567--80. \url{https://doi.org/10.1162/rest_a_00732}.

\leavevmode\hypertarget{ref-Athey2019}{}%
{[}14{]} Athey S, Tibshirani J, Wager S. Generalized random forests. The
Annals of Statistics 2019;47. \url{https://doi.org/10.1214/18-aos1709}.

\leavevmode\hypertarget{ref-Lu2018}{}%
{[}15{]} Lu M, Sadiq S, Feaster DJ, Ishwaran H. Estimating individual
treatment effect in observational data using random forest methods.
Journal of Computational and Graphical Statistics 2018;27:209--19.
\url{https://doi.org/10.1080/10618600.2017.1356325}.

\leavevmode\hypertarget{ref-Wager2018}{}%
{[}16{]} Wager S, Athey S. Estimation and inference of heterogeneous
treatment effects using random forests. Journal of the American
Statistical Association 2018;113:1228--42.
\url{https://doi.org/10.1080/01621459.2017.1319839}.
\end{cslreferences}

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

\noindent

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}