### 2021-11-12 submission/manuscript.rmd
update abstract

---
title: |
 Linear interaction of treatment with baseline risk was sufficient for predicting
 individualized benefit
abstract: |
 **Objective:** To compare different risk-based methods predicting
 individualized treatment effects in RCTs. **Study Design and
 Setting:** We simulated data using diverse assumptions for a baseline
 prognostic index of risk (PI), the shape of its interaction with
 treatment (none, linear, quadratic or non-monotonic) and the magnitude
 of treatment-related harms. In each sample we predicted absolute
 benefit using: models with a constant relative treatment effect;
 stratification in quarters of the PI; models including a linear
 interaction of treatment with the PI; models including an interaction
 of treatment with a restricted cubic spline (RCS) transformation of
 the PI; an adaptive approach using Akaike’s Information Criterion. We
 evaluated predictive performance using root mean squared error and
 measures of discrimination and calibration for benefit. Starting from
 a base scenario (sample size 4,250, treatment odds ratio 0.8, AUC of
 the PI 0.75), we varied the sample size, the treatment effect size,
 and the PI's AUC. **Results:** Models including a PI by treatment
 interaction performed best with smaller sample sizes and/or lower AUC
 of the PI. Otherwise, RCS (3 knots) models proved more robust for the
 majority of the deviation settings. The adaptive approach was unstable
 with smaller sample sizes. **Conclusion:** Depending on the setting,
 the linear interaction or the RCS (3 knots) model should be preferred.

authors:
  - name: Alexandros Rekkas
    department: Department of Medical Informatics
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
  - name: Peter R. Rijnbeek
    department: Department of Medical Informatics
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
  - name: Ewout W. Steyerberg
    department: Department of Biomedical Data Sciences
    affiliation: Leiden University Medical Center
    location: Leiden, The Netherlands
  - name: David van Klaveren
    department: Department of Public Health
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
keywords:
  - treatment effect heterogeneity
  - absolute benefit
  - prediction models
output: 
  rticles::arxiv_article:
    keep_tex: true
  bookdown::word_document2:
    reference_docx: reference.docx
geometry: margin=1.0in
date: false
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \usepackage{amsmath}
  - \doublespacing
  - \usepackage[pagewise, left]{lineno}
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \date{}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(SmoothHte)
library(rms)
library(here)

d <- function(x, decimals = 2) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}

knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
set.seed(19910930)
```

\doublespacing 
\linenumbers

# Introduction
Predictive approaches for assessing heterogeneity of treatment effects (HTE) aim
at the development of models predicting either individualized effects or which
of two (or more) treatments is better for an individual [@Varadhan2013]. In
prior work, we divided such methods in three broader categories based on the
reference class used for defining patient similarity when making individualized
predictions or recommendations [@Rekkas2020]. Risk-modeling approaches use
prediction of baseline risk as the reference; treatment effect modeling
approaches also model treatment-covariate interactions, in addition to risk
factors; optimal treatment regime approaches focus on developing treatment
assignment rules and therefore rely heavily on modeling treatment effect
modifiers.

Risk-modeling approaches to predictive HTE analyses provide a viable option in
the absence of well-established treatment effect modifiers [@Kent2019;
@PathEnE]. In simulations, modeling of effect modifiers, i.e.
treatment-covariate interactions, often led to miscalibrated predictions of
benefit, while risk-based methods proved quite robust [@vanKlaveren2019]. Most
often, risk-modeling approaches are carried out in two steps: first a risk
prediction model is developed externally or internally on the entire RCT
population, “blinded” to treatment; then the RCT population is stratified using
this prediction model to evaluate risk-based treatment effect variation
[@Kent2010]. However, even though estimates at the risk subgroup level may be
accurate, these estimates do not apply to individual patients, especially for
patients with predicted risk at the boundaries of the risk intervals. Hence, the
risk-stratified approach is useful for exploring and presenting HTE, but is not
useful for supporting treatment decisions for individual patients.

To individualize treatment effects, the recent PATH statement suggested various
risk-based models including a prognostic index of baseline risk (PI) and
treatment assignment [@Kent2019; @PathEnE]. We aimed to summarize and compare
different risk-based models for predicting individualized treatment effects. We
simulated RCT settings to compare the performance of these models under
different assumptions of the relationship between baseline risk and treatment.
We illustrated the different models by a case study of predicting individualized
effects of tissue plasminogen activator (tPA) versus streptokinase treatment in
patients with an acute myocardial infarction (MI).

# Methods

## Simulation scenarios

For each patient we generated 8 baseline covariates
$x_1,\dots,x_4\sim N(0, 1)$ and $x_5,\dots,x_8\sim B(1, 0.2)$. Treatment
was allocated using a 50:50 split. Outcomes for patients in the control arm were
generated from a logistic regression model including all baseline covariates. In
the base scenarios coefficient values were such, that the AUC of the
logistic regression model was $0.75$ and the event rate in the control arm was
$20\%$. Binary outcomes in the control arm were generated from Bernoulli variables
with true probabilities 
$P(y=1|X, t_x = 0) = \text{expit}(PI)=\frac{e^{PI}}{1+e^{PI}}$.

Outcomes in the treatment arm were generated using 3 base scenarios: absent
treatment effect (OR = 1), moderate treatment effect (OR = 0.8) and high
treatment effect (OR = 0.5). We started with simulating outcomes based on true
constant relative treatment effects for the 3 base scenarios. We then simulated
linear, quadratic and non-monotonic deviations from constant treatment effects
using:
$$lp_1 = \gamma_2(PI-c)^2 + \gamma_1(PI-c) + \gamma_0, $$
where $lp_1$ is the true linear predictor in the treatment arm, so that
$P(y=1|X, t_x=1) = \text{expit}(lp_1)$. Finally, we simulated scenarios where a
constant absolute harm is applied across all treated patients. In this case we
have $P(y=1|X,t_x=1) = \text{expit}(lp_1) + \text{harm}$.

The sample size for the base scenarios was set to 4,250 ($80\%$ power for the
detection of a marginal OR of 0.8). We evaluated the effect of smaller or larger
sample sizes of 1,063 and 17,000, respectively. We also evaluated the effect of
worse or better discriminative ability for risk, adjusting the baseline
covariate coefficients, such that the AUC of the regression model in the control
arm was 0.65 and 0.85 respectively.

Combining all these settings resulted in a simulation study of 648 scenarios
(exact settings in the supplementary material).

## Individualized risk-based benefit predictions

All methods assume that a risk prediction model is available to assign risk
predictions to individual patients. For the simulations we developed a
prediction model internally, using logistic regression including main effects
for all baseline covariates and treatment assignment. Risk predictions for
individual patients were based on treatment assignment to the control arm, that
is setting treatment assignment to 0.

A *stratified HTE method* has been suggested as an alternative to traditional
subgroup analyses. Patients are stratified into equally-sized risk strata---in
this case based on risk quartiles. Absolute treatment effects within risk strata
are estimated by the difference in event rate between patients in the control
arm and patients in the treated arm. We considered this approach as a reference,
expecting it to perform worse than the other candidates, as its objective is not
to individualize benefit prediction.

Second, we considered a model which assumes *constant relative treatment effect*
(constant odds ratio). Hence, absolute benefit is predicted from
$\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))$.

Third, we considered a logistic regression model including treatment, the
prognostic index, and their linear interaction. Absolute benefit is then
estimated from
$\hat{\tau}(\bm{x})=\text{expit}(\beta_0+\beta_{PI}PI) - \text{expit}(\beta_0+\beta_{t_x}+(\beta_{PI}+\beta_*)PI)$.
We will refer to this method as the *linear interaction* approach.

Fourth, we used *restricted cubic splines* (RCS) to relax the
linearity assumption on the effect of the linear predictor
[@Harrell1988]. We considered splines with 3 (RCS-3), 4 (RCS-4) and 5
(RCS-5) knots to compare models with different levels of flexibility.

Finally, we considered an adaptive approach using Akaike's Information Criterion
(AIC) for model selection. The candidate models were: a constant treatment
effect model, a model with a linear interaction with treatment and RCS models
with 3, 4 and 5 knots.

## Evaluation metrics
We evaluated the predictive accuracy of the considered methods by the root mean
squared error (RMSE):

$$\text{RMSE}=\sqrt{\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2}$$
We compared the discriminative ability of the methods under study using
c-for-benefit [@vanKlaveren2018]. The c-for-benefit represents the probability
that from two randomly chosen matched patient pairs with unequal observed
benefit, the pair with greater observed benefit also has a higher predicted
benefit. To be able to calculate observed benefit, patients in each treatment
arm are ranked based on their predicted benefit and then matched 1:1 across
treatment arms. *Observed* treatment benefit is defined as the difference of
observed outcomes between the untreated and the treated patient of each matched
patient pair. *Predicted* benefit is defined as the average of predicted benefit
within each matched patient pair.

We evaluated calibration in a similar manner, using the integrated calibration
index (ICI) for benefit [@Austin2019]. The observed benefits are regressed on
the predicted benefits using a locally weighted scatterplot smoother (loess).
The ICI-for-benefit is the average absolute difference between predicted and
smooth observed benefit. Values closer to $0$ represent better calibration.

# Results

## Simulations

```{r adaptive, echo=FALSE, warning=FALSE, message=FALSE}
adaptiveSelections <- readr::read_csv(here::here("data/processed/adaptiveSelections.csv"))
rmseDistribution <- readr::read_csv(here::here("data/processed/rmseDistribution.csv"))
aucDistribution <- readr::read_csv(here::here("data/processed/discriminationDistribution.csv"))
```
The linear interaction model outperformed all RCS methods in terms of
RMSE in scenarios with true constant relative treatment effect (OR =
0.8, N = `r 4250` and AUC = 0.75), strong linear and even strong
quadratic deviations (Figure \ref{fig:rmsebase}). However, with
non-monotonic deviations errors of the linear interaction model
increased substantially, especially in the presence of
treatment-related harms. In these scenarios, RCS-3 proved to be quite
robust outperforming all other methods. The constant treatment effect
approach had overall best performance under true constant treatment
effect settings, but was sensitive to all considered deviations,
resulting in increased error rates. Finally, the adaptive approach had
comparable performance to the best-performing method in each
scenario. However, we observed increased error variability in the case
of linear and non-monotonic deviations, especially for moderate or
strong treatment-related harms.

In the case of true constant relative treatment effects, the adaptive
approach selected the constnant effect model the majority of the time,
even in the presence of treatment-related harms
(`r adaptiveSelections %>% filter(scenarioId == 217) %>% pull(treatment) * 100`\%
of the time with no harms to
`r adaptiveSelections %>% filter(scenarioId == 219) %>% pull(treatment) * 100`\% 
with strong harms). With strong linear or quadratic deviations and
increasing treatment-related harms the adaptive approach increasingly
favored the linear interaction model. In the case of true
non-monotonic deviations, we observed increasing selection frequencies
for RCS-3 with increasing treatment harms (from
`r adaptiveSelections %>% filter(scenarioId == 397) %>% pull(rcs_3) * 100`\%
with no harms to
`r adaptiveSelections %>% filter(scenarioId == 399) %>% pull(rcs_3) * 100`\%
with strong harms), whereas selection frequencies for the linear
interaction model dropped from
`r adaptiveSelections %>% filter(scenarioId == 397) %>% pull(risk) * 100`\%
with no harms to
`r adaptiveSelections %>% filter(scenarioId == 399) %>% pull(risk) * 100`\%
with strong harms. Finally, in the case of strong linear and
non-monotonic deviations selection frequencies of the constant
treatment effect model remained high despite increasing treatment
harms (Supplement, Table XX).

```{r rmsebase, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated super-population of size 500,000. The scenario with true constant relative treatment effect had a true prediction AUC of 0.75 and sample size of 4250. Results are presented under moderate linear, strong linear, and strong quadratic deviations from constant relative treatment effects.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_base.tiff")))
# knitr::include_graphics(here::here("figures/rmse_base.png"))
```

Increasing the sample size to `r 17000` favored RCS-3 the most, as it
achieved lowest or close to lowest RMSE across all scenarios (Figure
\ref{fig:rmsesamplesize}). Especially in cases of strong quadratic and
non-monotonic deviations RCS-3 had lower error rates (median RMSE
`r rmseDistribution %>% filter(scenarioId == 385) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_rcs_3)`
for strong quadratic deviations and 
`r rmseDistribution %>% filter(scenarioId == 421) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_rcs_3)`
for non-monotonic deviations with no treatment-related harms) compared
to the linear interaction approach (median RMSE
`r rmseDistribution %>% filter(scenarioId == 385) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_linear)` and 
`r rmseDistribution %>% filter(scenarioId == 421) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_linear)`,
respectively), regardless of treatment-related harms strength. The
issues with the large error variability of the adaptive approach
improved with larger sample sizes.

The adaptive approach tended to increasingly favor smoother methods
(especially RCS-3) with increasing treatment-related harms (see
Supplement, Table XX). However, in the case of true strong quadratic
deviations the opposite was observed: selection frequency of the
linear interaction model increased from
`r adaptiveSelections %>% filter(scenarioId == 385) %>% pull(risk) * 100`\% (no harms)
to
`r adaptiveSelections %>% filter(scenarioId == 387) %>% pull(risk) * 100`\% (strong harms)
whereas for RCS-3 decreased from
`r adaptiveSelections %>% filter(scenarioId == 385) %>% pull(rcs_3) * 100`\% (no harms)
to
`r adaptiveSelections %>% filter(scenarioId == 387) %>% pull(rcs_3) * 100`\% (strong harms).

```{r rmsesamplesize, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. Sample size 17,000 rather than 4250 in Figure \\ref{fig:rmsebase}", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_sample_size.tiff")))
# knitr::include_graphics(here::here("figures/rmse_sample_size.png"))
```

When we increased the AUC of the true prediction model to 0.85 (OR =
0.8 and N = `r 4250`) the constant effect model outperformed all other
methods in the case of true constant treatment effects, but proved to
be the least robust to deviations. Again, RCS-3 had the lowest
error rates in the case of strong quadratic or monotonic deviations
and very comparable performance to the best-performing linear
interaction model in the case of strong linear deviations
(median RMSE
`r rmseDistribution %>% filter(scenarioId == 297) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_rcs_3)` for RCS-3 compared to 
`r rmseDistribution %>% filter(scenarioId == 297) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_linear)`
for the linear interaction model). The adaptive approach, though it
performed similar to the best performing method in each scenario, on
average, had increased variability in error rates in the case of
strong linear and non-monotonic deviations. In these scenarios, the
adaptive approach often selected the constant treatment effect model
(`r adaptiveSelections %>% filter(scenarioId == 297) %>% pull(treatment) * 100`\%
and
`r adaptiveSelections %>% filter(scenarioId == 405) %>% pull(treatment) * 100`\%
in the strong linear and non-monotonic deviation scenarios without
treatment-related harms, respectively).

```{r rmseauc, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.85 and sample size of 4,250.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_auc.tiff")))
# knitr::include_graphics(here::here("figures/rmse_auc.png"))
```

In terms of discrimination for benefit (OR = 0.8, N = `r 4250` and AUC
= 0.75), all methods performed similarly, on average (Figure
\ref{fig:discrimination}). In the case of non-monotonic deviations,
the constant effect model had much lower discriminative performance
compared to the rest of the methods (median AUC of
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_constant)` 
for the constant effects model compared to the best-performing RCS-3
with
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_rcs_3)`).
The linear interaction model was the most stable compared to the other
methods in terms of error variability. With increasing number of RCS
knots, we observed decreasing median values and increasing variability
of the c-for-benefit in all scenarios.

```{r discrimination, cache=TRUE, echo=FALSE, fig.cap="Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.75 and sample size of 4,250.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/discrimination_base.tiff")))
# knitr::include_graphics(here::here("figures/discrimination_base.png"))
```

In terms of calibration for benefit, the constant effects model
outperformed all other models in the case of true constant treatment
effects, but was miscalibrated for all deviation scenarios (Figure
\ref{fig:calibration}). The linear interaction model showed best or
close to best calibration across all scenarios and only showed worse
calibration compared to RCS-3 in the case of non-monotonic deviations
and treatment-related harms. The adaptive approach was worse
calibrated in scenarios with strong linear and non-monotonic
deviations compared to the linear interaction model and RCS-3.

The results from all individual scenarios can be explored online at
[https://arekkas.shinyapps.io/simulation_viewer/](https://arekkas.shinyapps.io/simulation_viewer/).

```{r calibration, cache=TRUE, echo=FALSE, fig.cap="Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.75 and sample size of 4,250.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/calibration_base.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

## Case study
```{r, echo=FALSE, warning=FALSE, message=FALSE}
load(here::here("data/raw/gusto.rda"))
gusto <- gusto %>%
  tibble() %>%
  filter(!is.na(tpa))

treatmentArms <- gusto %>%
  group_by(tpa) %>%
  summarise(n = n())
```

We demonstrate the different methods for individualizing treatment
benefits using data from `r nrow(gusto)` patients with an acute
myocardial infarction (MI) included in the GUSTO-I trial.
`r treatmentArms %>% filter(tpa == 1) %>% select(n)` patients were randomized to
tissue plasminogen activator (tPA) treatment and
`r treatmentArms %>% filter(tpa == 0) %>% select(n)` were randomized to
streptokinase. The outcome of interest was 30-day mortality, recorded
for all patients.

In line with previous analyses [@Califf1997; @Steyerberg2000], we
fitted a logistic regression model with 6 baseline covariates,
i.e. age, Killip class, systolic blood pressure, heart rate, an
indicator of previous MI, and the location of MI, to predict 30-day
mortality risk. A constant effect of treatment was included in the
model. When deriving risk predictions for individuals we set the
treatment indicator to 0. More information on model development can be
found in the supplement.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
gustoPerformance <- readr::read_csv(here::here("data/processed/gustoPerformanceMetrics.csv"))

discriminationPerformance <- gustoPerformance %>%
  select("discrimination")
calibrationPerformance <- gustoPerformance %>%
  select("calibration")
```

We used the risk linear predictor to fit the proposed methods under
study for individualizing absolute benefit predictions. All methods
predicted increasing benefits for patients with higher baseline risk
predictions, but the fitted patterns were clearly different. The
adaptive approach selected the model with RCS smoothing with 4
knots. However, for very low baseline risk this model predicted
decreasing benefit with increasing risk may be somewhat too flexible.
The more robust models, the linear interaction model or the model with
RCS smoothing (3 knots), gave very similar benefit predictions,
followed the evolution of the stratified estimates very closely and
may therefore be preferable for use in clinical practice. The linear
interaction model had somewhat lower AIC compared to the model with
RCS smoothing (3 knots), slightly better cross-validated
discrimination (c-for-benefit
`r round(discriminationPerformance[2, ], 3)` vs
`r round(discriminationPerformance[3, ], 3)`) and quite similar cross-validated
calibration (ICI-for benefit `r round(calibrationPerformance[2, ], 4)` vs
`r round(calibrationPerformance[3, ], 4)`).

```{r gusto, cache=TRUE, echo=FALSE, fig.cap="Individualized absolute benefit predictions based on baseline risk when using a constant treatment effect approach, a linear interaction approach and RCS smoothing using 3,4 and 5 knots. Risk stratified estimates of absolute benefit are presented within quartiles of baseline risk as reference.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/gusto.png"))
```

# Discussion

The linear interaction model displayed very good performance overall
under many of the considered simulation scenarios. Especially in cases
with smaller sample sizes and moderately performing baseline risk
prediction models it had lower RMSE, was better calibrated for benefit
and had better discrimination for benefit, even in scenarios with
strong quadratic deviations. However, in scenarios with true
non-monotonic deviations, the linear interaction model was
outperformed by RCS-3, especially in the presence of true
treatment-related harms. Increasing the sample size or the prediction
model's discrimination favored RCS-3 which had better or very
comparable performance to the linear interaction model, but was more
robust to non-monotonic deviations and the presence of
treatment-related harms.

RCS-4 and RCS-5 proved to be too flexible, as indicated by higher
RMSE, increased variability of discrimination for benefit and worse
calibration of benefit predictions. Even with larger sample sizes and
strong quadratic or non-monotonic deviations from the base case
scenario of constant relative treatment effects, these more flexible
restricted cubic splines did not outperform the simpler RCS-3 These
approaches may only be helpful if we expect more extreme patterns of
heterogeneous treatment effects compared to the quadratic deviations
considered here.

The constant treatment effect model, despite having adequate
performance in the presence of weak treatment effect heterogeneity on
the relative scale, quickly broke down with stronger deviations from
constant relative treatment effects. In these cases, the stratified
approach generally had lower error rates compared to the constant
treatment effect model. Stepwise treatment benefit estimates are very
useful for demonstrating treatment effect heterogeneity--because
estimating treatment effect requires groups of patients rather than
individual patients--but are not helpful for making individualized
absolute benefit predictions.

Increasing the discriminative ability of the risk model--by increasing
the predictor coefficients of the true risk model--reduced RMSE for
all methods.  This increase in discriminative ability translates in
higher variability of predicted risks, which, in turn, allows the
considered methods to better capture absolute treatment benefits. As a
consequence, the increase in discriminative ability of the risk model
also led to higher values of c-for-benefit. Even though risk model
performance is very important for the ability of risk-based methods to
predict treatment benefit, prediction model development was outside
the scope of this work and has already been studied extensively
[@vanKlaveren2019; @Burke2014; @Abadie2018].

The adaptive approach had adequate performance, following closely on
average the performance of the "true" model in most
scenarios. However, with smaller sample sizes it tended to "miss" the
treatment-risk interactions and selected simpler models (Supplementary
Table S7). This resulted in increased RMSE variability in these
scenarios, especially in the case of true strong linear or
non-monotonic deviations from the base case scenario.  Therefore, in
the case of smaller sample sizes the simpler linear interaction model
is a safer choice for predicting absolute benefits.

Risk-based approaches to predictive HTE estimate treatment benefit as
a function of baseline risk. A limitation of our study is that we
assumed treatment benefit to be a function of baseline risk in the
majority of the simulation scenarios. We attempted to address that by
introducing constant moderate and strong treatment-related harms,
applied on the absolute scale. Also, we considered a small number of
scenarios with true treatment-covariate interactions, in which our
main conclusions remained the same (Supplement, XX). Future simulation
studies could explore the effect of more extensive deviations from
risk-based treatment effects.

Recent years have seen an increased interest in predictive HTE
approaches focusing on individualized benefit predictions. In our
simulations we only focused on risk-based methods, using baseline risk
as a reference in a two-stage approach to individualizing benefit
predictions. However, there is a plethora of different methods,
ranging from treatment effect modeling to tree-based approaches
available in more recent literature [@Athey2019; @Lu2018;
@Wager2018]. Simulations are also needed to assess relative
performance and define the settings where these break down or
outperform each other.

In conclusion, the best option for predicting individualized treatment
benefit using a risk-based approach depends on the setting. With
smaller sample sizes and/or moderately performing risk prediction
models the linear interaction approach is a viable option. When those
constraints are not present or when we anticipate non-negligible
treatment-related harms, RCS-3 is a better option in terms of error
rates, discrimination and calibration for benefit. With larger sample
size, an adaptive approach based on AIC can also be considered as a
more automated alternative.

\newpage
# References
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent