### 2021-12-15 submission/manuscript.rmd
shared version

---
title: |
  An interaction between baseline risk and treatment assignment 
  generally imporoved treatment effect predictions
abstract: |
 **Objective**: To compare different risk-based methods predicting
 individualized treatment effects in RCTs. **Study Design and
 Setting**: We simulated RCT data using diverse assumptions for the
 average treatment effect, a baseline prognostic index of risk (PI),
 the shape of its interaction with treatment (none, linear, quadratic
 or non-monotonic) and the magnitude of treatment-related harms. In
 each sample we predicted absolute benefit using: models with a
 constant relative treatment effect; stratification in quarters of the
 PI; models including a linear interaction of treatment with the PI;
 models including an interaction of treatment with a restricted cubic
 spline (RCS) transformation of the PI; an adaptive approach using
 Akaike’s Information Criterion. We evaluated predictive performance
 using root mean squared error and measures of discrimination and
 calibration for benefit. **Results**: The linear-interaction model
 and the RCS-interaction model outperformed the constant treatment
 effect model in many simulation scenarios. The RCS-model was optimal
 when quadratic or non-monotonic deviations from a constant treatment
 effect were stronger, and when sample size was larger. Larger sample
 size also supported the adaptive approach. **Conclusion**: An
 interaction between risk and treatment assignment generally improved
 treatment effect predictions. Non-linear RCS interactions should be
 considered for larger sample size.
author:
  - name: Alexandros Rekkas
    affiliation: 1
  - name: Peter R. Rijnbeek
    affiliation: 1
  - name: David M. Kent
    affiliation: 2
  - name: Ewout W. Steyerberg
    affiliation: 3
  - name: David van Klaveren
    affiliation: 4
address:
  - code: 1
    address: Department of Medical Informatics, Erasmus Medical Center, Rotterdam, The Netherlands
  - code: 2
    address: Predictive Analytics and Comparative Effectiveness Center, Institute for Clinical Research and Health Policy Studies, Tufts Medical Center, Boston, Massachusetts, USA
  - code: 3
    address: Department of Biomedical Data Sciences, Leiden University Medical Center, Leiden, The Netherlands
  - code: 4
    address: Department of Public Health, Erasmus Medical Center, Rotterdam, The Netherlands
numbersections: true
linenumbers: false
journal: arXiv.org
keywords:
  - treatment effect heterogeneity
  - absolute benefit
  - prediction models
output: 
  rticles::elsevier_article:
    keep_tex: true
  bookdown::word_document2:
    reference_docx: reference.docx
geometry: margin=1.0in
date: false
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \usepackage{amsmath}
  - \doublespacing
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \date{}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(SmoothHte)
library(rms)
library(here)

d <- function(x, decimals = 4) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}

knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
set.seed(19910930)
```

\doublespacing 
\linenumbers

# Introduction
Predictive approaches for assessing heterogeneity of treatment effects (HTE) aim
at the development of models predicting either individualized effects or which
of two (or more) treatments is better for an individual [@Varadhan2013]. In
prior work, we divided such methods in three broader categories based on the
reference class used for defining patient similarity when making individualized
predictions or recommendations [@Rekkas2020]. Risk-modeling approaches use
prediction of baseline risk as the reference; treatment effect modeling
approaches also model treatment-covariate interactions, in addition to risk
factors; optimal treatment regime approaches focus on developing treatment
assignment rules and therefore rely heavily on modeling treatment effect
modifiers.

Risk-modeling approaches to predictive HTE analyses provide a viable option in
the absence of well-established treatment effect modifiers [@Kent2019;
@PathEnE]. In simulations, modeling of effect modifiers, i.e.
treatment-covariate interactions, often led to miscalibrated predictions of
benefit, while risk-based methods proved quite robust [@vanKlaveren2019]. Most
often, risk-modeling approaches are carried out in two steps: first a risk
prediction model is developed externally or internally on the entire RCT
population, “blinded” to treatment; then the RCT population is stratified using
this prediction model to evaluate risk-based treatment effect variation
[@Kent2010]. This two-step approach identified substantial absolute treatment
effect differences between low-risk and high-risk patients in a re-analysis of
32 large trials [@Kent2016]. However, even though estimates at the risk subgroup
level may be accurate, these estimates do not apply to individual patients,
especially for patients with predicted risk at the boundaries of the risk
intervals. Hence, the risk-stratified approach is useful for exploring and
presenting HTE, but is not useful for supporting treatment decisions for
individual patients.

To individualize treatment effects, the recent PATH statement suggested various
risk-based models including a prognostic index of baseline risk (PI) and
treatment assignment [@Kent2019; @PathEnE]. We aimed to summarize and compare
different risk-based models for predicting individualized treatment effects. We
simulated RCT settings to compare the performance of these models under
different assumptions of the relationship between baseline risk and treatment.
We illustrated the different models by a case study of predicting individualized
effects of tissue plasminogen activator (tPA) versus streptokinase treatment in
patients with an acute myocardial infarction (MI).

# Methods

## Simulation scenarios

We simulated a typical RCT that is undertaken to compare a binary outcome
(e.g. death) between a group of treated patients in the treatment arm and a
group of untreated patients in the control arm. For each patient we generated 8
baseline covariates $x_1,\dots,x_4\sim N(0, 1)$ and $x_5,\dots,x_8\sim B(1,
0.2)$. Treatment was allocated using a 50:50 split. Outcomes for patients in the
control arm were generated from a logistic regression model including all
baseline covariates. In the base scenarios coefficient values were such, that
the AUC of the logistic regression model was $0.75$ and the event rate in the
control arm was $20\%$. Binary outcomes in the control arm were generated from
Bernoulli variables with true probabilities
$P(y=1|X, t_x = 0) = \text{expit}(PI)=\frac{e^{PI}}{1+e^{PI}}$.

Outcomes in the treatment arm were generated using 3 base scenarios: absent
treatment effect (OR = 1), moderate treatment effect (OR = 0.8) and strong
treatment effect (OR = 0.5). We started with simulating outcomes based on true
constant relative treatment effects for the 3 base scenarios. We then simulated
linear, quadratic and non-monotonic deviations from constant treatment effects
using:
$$lp_1 = \gamma_2(PI-c)^2 + \gamma_1(PI-c) + \gamma_0, $$
where $lp_1$ is the true linear predictor in the treatment arm, so that
$P(y=1|X, t_x=1) = \text{expit}(lp_1)$. Finally, we simulated scenarios where a
constant absolute harm is applied across all treated patients. In this case we
have $P(y=1|X,t_x=1) = \text{expit}(lp_1) + \text{harm}$.

The sample size for the base scenarios was set to 4,250 ($80\%$ power for the
detection of a marginal OR of 0.8). We evaluated the effect of smaller or larger
sample sizes of 1,063 (4,250 divided by 4) and 17,000 (4250 multiplied by 4),
respectively. We also evaluated the effect of worse or better discriminative
ability for risk, adjusting the baseline covariate coefficients, such that the
AUC of the regression model in the control arm was 0.65 and 0.85 respectively.

Combining all these settings resulted in a simulation study of 648 scenarios
(exact settings in the supplementary material). With these scenarios we were
able to cover the observed treatment effect heterogeneity in 32 large trials as
well as many other potential variations of risk-based treatment effect
[@Kent2016].

## Individualized risk-based benefit predictions

All risk-based methods assume that a risk prediction model is available to
assign risk predictions to individual patients. For the simulations we developed
a prediction model internally, using logistic regression including main effects
for all baseline covariates and treatment assignment. Risk predictions for
individual patients were based on treatment assignment to the control arm, that
is setting treatment assignment to 0.

A *stratified HTE method* has been suggested as an alternative to traditional
subgroup analyses. Patients are stratified into equally-sized risk strata---in
this case based on risk quartiles. Absolute treatment effects within risk strata
are estimated by the difference in event rate between patients in the control
arm and patients in the treated arm. We considered this approach as a reference,
expecting it to perform worse than the other candidates, as its objective is not
to individualize benefit prediction.

Second, we considered a model which assumes *constant relative treatment effect*
(constant odds ratio). Hence, absolute benefit is predicted from
$\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))$.

Third, we considered a logistic regression model including treatment, the
prognostic index, and their linear interaction. Absolute benefit is then
estimated from
$\hat{\tau}(\bm{x})=\text{expit}(\beta_0+\beta_{PI}PI) - \text{expit}(\beta_0+\beta_{t_x}+(\beta_{PI}+\beta_*)PI)$.
We will refer to this method as the *linear interaction* approach.

Fourth, we used *restricted cubic splines* (RCS) to relax the linearity
assumption on the effect of the linear predictor [@Harrell1988]. We considered
splines with 3 (RCS-3), 4 (RCS-4) and 5 (RCS-5) knots to compare models with
different levels of flexibility.

Finally, we considered an adaptive approach using Akaike's Information Criterion
(AIC) for model selection. The candidate models were: a constant treatment
effect model, a model with a linear interaction with treatment and RCS models
with 3, 4 and 5 knots.

## Evaluation metrics
We evaluated the predictive accuracy of the considered methods by the root mean
squared error (RMSE):

$$\text{RMSE}=\sqrt{\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2}$$
We compared the discriminative ability of the methods under study using
c-for-benefit [@vanKlaveren2018]. The c-for-benefit represents the probability
that from two randomly chosen matched patient pairs with unequal observed
benefit, the pair with greater observed benefit also has a higher predicted
benefit. To be able to calculate observed benefit, patients in each treatment
arm are ranked based on their predicted benefit and then matched 1:1 across
treatment arms. *Observed* treatment benefit is defined as the difference of
observed outcomes between the untreated and the treated patient of each matched
patient pair. *Predicted* benefit is defined as the average of predicted benefit
within each matched patient pair.

We evaluated calibration in a similar manner, using the integrated calibration
index (ICI) for benefit [@Austin2019]. The observed benefits are regressed on
the predicted benefits using a locally weighted scatterplot smoother (loess).
The ICI-for-benefit is the average absolute difference between predicted and
smooth observed benefit. Values closer to $0$ represent better calibration.

# Results

## Simulations

```{r adaptive, echo=FALSE, warning=FALSE, message=FALSE}
adaptiveSelections <- readr::read_csv(here::here("data/processed/adaptiveSelections.csv"))
rmseDistribution <- readr::read_csv(here::here("data/processed/rmseDistribution.csv"))
aucDistribution <- readr::read_csv(here::here("data/processed/discriminationDistribution.csv"))
```

The linear interaction model outperformed all RCS methods in terms of
RMSE in scenarios with true constant relative treatment effect (OR =
0.8, N = `r 4250` and AUC = 0.75), strong linear and even strong
quadratic deviations from a constant relative treatment effect (Figure
\ref{fig:rmsebase}; panels A-C). However, with non-monotonic
deviations from a constant relative treatment effect, the RMSE of the
linear interaction model increased substantially, especially in the
presence of treatment-related harms (Figure \ref{fig:rmsebase}; panel
D). In these scenarios, RCS-3 outperformed all other methods in terms
of RMSE. The constant treatment effect approach had overall best
performance under true constant treatment effect settings, but was
sensitive to all considered deviations, resulting in increased
RMSE. Finally, the adaptive approach had comparable performance to the
best-performing method in each scenario. However, in comparison with
the best-performing approach, its RMSE was more variable across the
500 replications in the scenarios with linear and non-monotonic
deviations, especially when also including moderate or strong
treatment-related harms. This is caused by wrongly selecting the
constant treatment effect model in a substantial proportion of the
replications (Supplement, Figure S1).

```{r rmsebase, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated super-population of size 500,000. The scenario with true constant relative treatment effect (panel A) had a true prediction AUC of 0.75 and sample size of 4250. The RMSE is also presented for strong linear (panel B), strong quadratic (panel C), and non-monotonic (panel D) from constant relative treatment effects. Panels on the right side present the true relationship between baseline risk (x-axis) and absolute treatment benefit (y-axis). The 2.5, 25, 50, 75, and 97.5 percentiles of the risk distribution are expressed by the boxplot.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_moderate_base.tiff")))
# knitr::include_graphics(here::here("figures/rmse_base.png"))
```

Increasing the sample size to `r 17000` favored RCS-3 the most, as it
achieved lowest or close to lowest RMSE across all scenarios (Figure
\ref{fig:rmsesamplesize}). Especially in cases of strong quadratic and
non-monotonic deviations RCS-3 had lower RMSE (median 
`r rmseDistribution %>% filter(scenarioId == 385) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_rcs_3)`
for strong quadratic deviations and 
`r rmseDistribution %>% filter(scenarioId == 421) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_rcs_3)`
for non-monotonic deviations with no treatment-related harms) compared
to the linear interaction approach (median 
`r rmseDistribution %>% filter(scenarioId == 385) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_linear)` and 
`r rmseDistribution %>% filter(scenarioId == 421) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_linear)`,
respectively), regardless of the strength of treatment-related
harms. Due to the large sample size, the RMSE of the adaptive approach
was even more similar to the best-performing method, and the constant
relative treatment effect model was less often wrongly selected
(Supplement, Figure S2).

```{r rmsesamplesize, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. Sample size 17,000 rather than 4250 in Figure \\ref{fig:rmsebase}", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_moderate_sample_size.tiff")))
# knitr::include_graphics(here::here("figures/rmse_sample_size.png"))
```

When we increased the AUC of the true prediction model to 0.85 (OR =
0.8 and N = `r 4250`). RCS-3 had the lowest RMSE in the case of strong
quadratic or non-monotonic deviations and very comparable performance to
the -- optimal -- linear interaction model in the case of strong
linear deviations (median RMSE
`r rmseDistribution %>% filter(scenarioId == 297) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_rcs_3)` for RCS-3 compared to 
`r rmseDistribution %>% filter(scenarioId == 297) %>% mutate(across(-one_of("scenarioId"), ~ d(.x, decimals = 3))) %>% pull(median_linear)`
for the linear interaction model). As observed in the base case
scenario the adaptive approach wrongly selected the constant treatment
effect model
(`r adaptiveSelections %>% filter(scenarioId == 297) %>% pull(treatment) * 100`\%
and
`r adaptiveSelections %>% filter(scenarioId == 405) %>% pull(treatment) * 100`\%
of the replications in the strong linear and non-monotonic deviation
scenarios without treatment-related harms, respectively), leading to
more variability of the RMSE (Supplement, Figure S3).

```{r rmseauc, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.85 and sample size of 4,250.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_moderate_auc.tiff")))
# knitr::include_graphics(here::here("figures/rmse_auc.png"))
```

In comparison with the true approach, dsicrimination for benefit in
the scenario with a constant relative treatment effect was only
slightly lower for the linear interaction model, but substantially
lower for the non-linear RCS approaches (Figure
\ref{fig:discrimination}; panel A). With strong linear or quadratic
deviations from a constant relative treatment effect, all methods
discriminated quite similarly (Figure \ref{fig:discrimination}; panels
B-C). In the scenario with non-monotonic deviations, the constant
effect model had much lower discriminative ability compared to
all other methods (median AUC of
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_constant)` 
for the constant effects model,
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_linear)`
for the linear interaction model and
`r aucDistribution %>% filter(scenarioId == 397) %>% mutate(across(-one_of("scenarioId"), ~ d(.x))) %>% pull(median_rcs_3)`
for the best-performing RCS-3; Figure \ref{fig:discrimination}; panel
D). The adaptive approach was unstable in terms of discrimination for
benefit, especially in the presence of treamtnt-related harms. With
increasing number of RCS knots, we observed decreasing median values
and increasing variability of the c-for-benefit in all scenarios. When
we increased the sample size to 17,000 we observed similar trends,
however the performance of all methods was more stable (Supplement,
Figure S4). Finally, when we increased the true prediction AUC to 0.85
the adaptive approach in the case of non-monotonic deviations was,
again, quite unstable, especially with null or moderate
treatment-related harms (Supplement, Figure S5). In these scenarios,
the adaptive approach tended to select more often the much inferior
constant treatment effect method (Supplement, Figure S3)

```{r discrimination, cache=TRUE, echo=FALSE, fig.cap="Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.75 and sample size of 4,250.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/discrimination_moderate_base.tiff")))
# knitr::include_graphics(here::here("figures/discrimination_base.png"))
```

In terms of calibration for benefit, the constant effects model
outperformed all other models in the scenario with true constant
treatment effects, but was miscalibrated for all deviation scenarios
(Figure \ref{fig:calibration}). The linear interaction model showed
best or close to best calibration across all scenarios and only showed
worse calibration compared to RCS-3 in case of non-monotonic
deviations and treatment-related harms (Figure \ref{fig:calibration};
panel D). The adaptive approach was worse calibrated in scenarios with
strong linear and non-monotonic deviations compared to the linear
interaction model and RCS-3. When we increased sample size to 17,000
similar conclusions on calibration for benefit could be drawn. As
expected, all methods displayed more stable calibration performance
due to the larger number of patients (Supplement, Figure S6). When we
increased the true prediction AUC to 0.85, the linear interaction
model was worse calibrated, on average, than RCS-3 in the case of
strong quadratic deviations from constant relative treatment effects
(Supplement, Figure S7).

The results from all individual scenarios can be explored online at
[https://arekkas.shinyapps.io/simulation_viewer/](https://arekkas.shinyapps.io/simulation_viewer/).

```{r calibration, cache=TRUE, echo=FALSE, fig.cap="Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.75 and sample size of 4,250.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/calibration_moderate_base.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

## Case study
```{r, echo=FALSE, warning=FALSE, message=FALSE}
load(here::here("data/raw/gusto.rda"))
gusto <- gusto %>%
  tibble() %>%
  filter(!is.na(tpa))

treatmentArms <- gusto %>%
  group_by(tpa) %>%
  summarise(n = n())
```

We demonstrate the different methods for individualizing treatment
benefits using data from `r nrow(gusto)` patients with an acute
myocardial infarction (MI) included in the GUSTO-I trial.
`r treatmentArms %>% filter(tpa == 1) %>% select(n)` patients were randomized to
tissue plasminogen activator (tPA) treatment and
`r treatmentArms %>% filter(tpa == 0) %>% select(n)` were randomized to
streptokinase. The outcome of interest was 30-day mortality, recorded
for all patients.

In line with previous analyses [@Califf1997; @Steyerberg2000], we
fitted a logistic regression model with 6 baseline covariates,
i.e. age, Killip class, systolic blood pressure, heart rate, an
indicator of previous MI, and the location of MI, to predict 30-day
mortality risk. A constant effect of treatment was included in the
model. When deriving risk predictions for individuals we set the
treatment indicator to 0. More information on model development can be
found in the supplement.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
gustoPerformance <- readr::read_csv(here::here("data/processed/gustoPerformanceMetrics.csv"))

discriminationPerformance <- gustoPerformance %>%
  select("discrimination")
calibrationPerformance <- gustoPerformance %>%
  select("calibration")
```

We used the risk linear predictor to fit the proposed methods under
study for individualizing absolute benefit predictions. All methods
predicted increasing benefits for patients with higher baseline risk
predictions, but the fitted patterns were clearly different. The
adaptive approach selected the model with RCS smoothing with 4
knots. However, for very low baseline risk the decreasing predicted
benefit with increasing risk may be somewhat too flexible. The more
robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed
the evolution of the stratified estimates very closely and may
therefore be preferable for use in clinical practice. The linear
interaction model had somewhat lower AIC compared to the model with
RCS smoothing (3 knots), slightly better cross-validated
discrimination (c-for-benefit
`r round(discriminationPerformance[2, ], 3)` vs
`r round(discriminationPerformance[3, ], 3)`) and quite similar cross-validated
calibration (ICI-for benefit `r round(calibrationPerformance[2, ], 4)` vs
`r round(calibrationPerformance[3, ], 4)`).

```{r gusto, cache=TRUE, echo=FALSE, fig.cap="Individualized absolute benefit predictions based on baseline risk when using a constant treatment effect approach, a linear interaction approach and RCS smoothing using 3,4 and 5 knots. Risk stratified estimates of absolute benefit are presented within quartiles of baseline risk as reference.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/gusto.tiff")))
# knitr::include_graphics(here::here("figures/gusto.png"))
```

# Discussion

The linear interaction model and the RCS-3 model both displayed very
good performance under many of the considered simulation scenarios, in
contrast with the constant relative treatment effect model. The linear
interaction model was optimal in cases with smaller sample sizes and
moderately performing baseline risk prediction models, that is, it had
lower RMSE, was better calibrated for benefit and had better
discrimination for benefit, even in scenarios with strong quadratic
deviations. In scenarios with true non-monotonic deviations, the
linear interaction model was outperformed by RCS-3, especially in the
presence of true treatment-related harms. Increasing the sample size
or the prediction model’s discriminative ability favored RCS-3,
especially in scenarios with non-monotonic deviations and in the
presence of treatment-related harms.

RCS-4 and RCS-5 proved to be too flexible in all considered scenarios,
as indicated by higher RMSE, increased variability of discrimination
for benefit and worse calibration of benefit predictions. Even with
larger sample sizes and strong quadratic or non-monotonic deviations
from the base case scenario of constant relative treatment effects,
these more flexible restricted cubic splines did not outperform the
simpler RCS-3 These approaches may only be helpful if we expect more
extreme patterns of heterogeneous treatment effects compared to the
quadratic deviations considered here.

The constant treatment effect model, despite having adequate
performance in the presence of weak treatment effect heterogeneity on
the relative scale, quickly broke down with stronger deviations from
constant relative treatment effects. In these cases, the stratified
approach generally had lower error rates compared to the constant
treatment effect model. Stepwise treatment benefit estimates are very
useful for demonstrating treatment effect heterogeneity–because
estimating treatment effect requires groups of patients rather than
individual patients–but are not helpful for making individualized
absolute benefit predictions.

Increasing the discriminative ability of the risk model–by increasing
the predictor coefficients of the true risk model–reduced RMSE for all
methods. This increase in discriminative ability translates in higher
variability of predicted risks, which, in turn, allows the considered
methods to better capture absolute treatment benefits. As a
consequence, the increase in discriminative ability of the risk model
also led to higher values of c-for-benefit. Even though risk model
performance is very important for the ability of risk-based methods to
predict treatment benefit, prediction model development was outside
the scope of this work and has already been studied extensively
[@vanKlaveren2019; @Burke2014; @Abadie2018].

The adaptive approach had adequate performance, following closely on
average the performance of the “true” model in most
scenarios. However, with smaller sample sizes it tended to “miss” the
treatment-risk interactions and selected simpler models (Supplementary
Table S7). This resulted in increased RMSE variability in these
scenarios, especially in the case of true strong linear or
non-monotonic deviations from the base case scenario. Therefore, in
the case of smaller sample sizes the simpler linear interaction model
is a safer choice for predicting absolute benefits.

Risk-based approaches to predictive HTE estimate treatment benefit as
a function of baseline risk. A limitation of our study is that we
assumed treatment benefit to be a function of baseline risk in the
majority of the simulation scenarios. We attempted to address that by
introducing constant moderate and strong treatment-related harms,
applied on the absolute scale. Also, we considered a small number of
scenarios with true treatment-covariate interactions, in which our
main conclusions remained the same (Supplement, XX). Future simulation
studies could explore the effect of more extensive deviations from
risk-based treatment effects.

Recent years have seen an increased interest in predictive HTE
approaches focusing on individualized benefit predictions. In our
simulations we only focused on risk-based methods, using baseline risk
as a reference in a two-stage approach to individualizing benefit
predictions. However, there is a plethora of different methods,
ranging from treatment effect modeling to tree-based approaches
available in more recent literature [@Athey2019; @Lu2018;@Wager2018].
Simulations are also needed to assess relative performance and define
the settings where these break down or outperform each other.

In conclusion, the linear interaction approach is a viable option with
smaller sample sizes and/or moderately performing risk prediction
modelsRCS-3 is a better option when non-monotonic deviations from a
constant relative treatment effect and/or substantial
treatment-related harms are anticipated. With larger sample size, an
adaptive approach -- selecting the method with optimal AIC -- can also
be considered as an automated alternative.

\newpage
# References
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent