### 2022-02-08 submission/supplement.rmd

---
title: "Supplementary material"
output:
    bookdown::pdf_document2: default
geometry: margin=1.0in
toc: true
font-size: 12pt
header-includes:
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

```{r, echo=FALSE, message=FALSE, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)
library(rms)

d <- function(x, decimals = 2) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}
knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      format = "f",
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
)
```

\beginsupplement
\newpage

# Simulation settings
For all patients we observe covariates $x_1,\dots,x_8$, of which $4$ are
continuous and $4$ are binary. More specifically,

\begin{equation*}
x_1,\dots,x_4 \sim N(0, 1)
\end{equation*}
\begin{equation*}
x_5,\dots,x_8 \sim B(1, 0.2)
\end{equation*}

We first, generate the binary outcomes $y$ for the untreated patients ($t_x=0$),
based on

\begin{equation} 
P(y\given \bm{x}, t_x=0) = g(\beta_0 + \beta_1x_1+\dots+\beta_8x_8) = g(lp_0),
(\#eq:p0)
\end{equation}

where $$g(x) = \frac{e^x}{1+e^x}$$

For treated patients, outcomes are generated from:

\begin{equation}
P(y\given\bm{x}, t_x=1) = g(lp_1)
(\#eq:p1)
\end{equation}

where $$lp_1 = \gamma_2(lp_0-c)^2+\gamma_1(lp_0-c)+\gamma_0$$

## Base-case scenario
The base-case scenario assumes a constant odds ratio of $0.8$ in favor of
treatment. The simulated datasets are of size $n=4250$, where treatment is
allocated at random using a 50/50 split (80% power for the detection of an
unadjusted OR of 0.8, assuming an event rate of 20% in the untreated
arm). Outcome incidence in the untreated population is set at $20\%$. For the
development of the prediction model we use the model defined in \@ref(eq:p0)
including a constant treatment effect. When doing predictions, $t_x$ is set to
$0$. The value of the true $\bm{\beta}$ is such that the above prediction model
has an AUC of $0.75$.

The previously defined targets are achieved when $\bm{\beta}=(-2.08,
0.49,\dots,0.49)^t$. For the derivations in the treatment arm we use
$\bm{\gamma}=(\log(0.8), 1, 0)^t$.

## Deviations from base-case
We deviate from the base-case scenario in two ways. First, we alter the overall
target settings of sample size, overall treatment effect and prediction model
AUC. In a second stage, we consider settings that violate the assumption of a
constant relative treatment effect, using a model-based approach.

For the first part, we consider:

* Sample size:
  + $n=1064$
  + $n=17000$
* Overall treatment effect:
  + $OR=0.5$
  + $OR=1$
* Prediction performance:
  + $AUC=0.65$
  + $AUC=0.85$

We set the true risk model coefficients to be
$\bm{\beta} = \big(-1.63,0.26,\dots,0.26\big)^t$ for $AUC=0.65$ and $\bm{\beta} = \big(-2.7,0.82,\dots,0.82\big)^t$ 
for $AUC=0.85$. In both cases, $\beta_0$ is selected so that an event rate of
$20\%$ is maintained in the control arm.

For the second part linear, quadratic and non-monotonic deviations from the
assumption of constant relative effect are considered. We also consider
different intensity levels of these deviations. Finally, constant absolute
treatment-related harms are introduced, i.e. positive 
($0.25\times\text{true average benefit}$),
strong positive ($0.50\times\text{true average benefit}$) and negative
($-0.25\times\text{true average benefit}$; i.e. constant absolute
treatment-related benefit). In case of true absent treatment effects,
treatment-related harms are set to $1\%, 2\%$ and $-1\%$ for positive, strong
positive and negative setting, respectively. The settings for these deviations
are defined in Table \@ref(tab:tab1).

```{r tab1, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
asd <- read_csv(
  here("data/processed/analysisIds.csv")
) %>%
  select(-"type")
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable   = TRUE,
  align       = "r",
  booktabs    = TRUE,
  caption     = "Scenario settings of the entire simulation study.",
  col.names   = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    paste0("g", 0:2),
    "c",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 3,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 4,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 7)
```

\newpage
# Plausible scenario settings

In this section we present specific scenarios from our simulation settings
in which evolution of benefit followed similar patterns to (REF!!). In this case
patients were stratified into risk quarters based on their true baseline
risk. Within each risk quarter we constructed boxplots of true benefit.

```{r, cache=TRUE, echo=FALSE, fig.cap="Simulation scenarios that closely follow trials. In this case, we see increasing absolute benefits with increasing baseline risk.", fig.show="hold", out.width = '50%'}
include_graphics(
  c(
    here("figures/scenario_251.png"),
    here("figures/scenario_324.png")
  )
)
```

```{r, cache=TRUE, echo=FALSE, fig.cap="Simulation scenarios that closely follow trials. In this case, we see increasing absolute benefits with increasing baseline risk.", fig.show="hold", out.width = '50%'}
include_graphics(
  c(
    here("figures/scenario_406.png"),
    here("figures/scenario_422.png")
  )
)
```

# Approaches to individualize benefit predictions

## Risk modeling
Merging treatment arms, we develop prediction models including a constant relative treatment effect:

\begin{equation}
P(y=1\given\bm{x}, t_x) = g(\bm{x}^t\bm{\beta} + \gamma t_x)
\end{equation}
(\#eq:risk)

The risk linear predictor is then 
$lp(\bm{x}; \bm{\beta}) = \bm{x}^t\bm{\beta}$. 
We derive baseline risk predictions for patients by setting $t_x=0$ in
\@ref(eq:risk). All methods for individualizing benefit predictions are 2-stage
methods, that start by fitting a model for predicting baseline risk. The
estimated linear predictor of this model is

\begin{equation*}
\hat{lp} = lp(\bm{x};\hat{\bm{\beta}}) = \bm{x}^t\hat{\bm{\beta}}
\end{equation*}

## Risk stratification
Derive a prediction model using the same approach as above and divide the
population in equally sized risk-based subgroups. Estimate subgroup-specific
absolute benefit from the observed absolute differences. Subject-specific
benefit predictions are made by attributing to individuals their corresponding
subgroup-specific estimate.

## Constant treatment effect
Assuming a constant relative treatment effect, fit the adjusted model in
\@ref(eq:risk). Then, an estimate of absolute benefit can be derived from
$$\tau(lp;  \hat{\bm{\beta}}, \hat{\gamma}) = g(\hat{lp}) - g(\hat{lp}+\hat{\gamma}) $$

## Linear interaction
The assumption of constant relative treatment effect is relaxed modeling a
linear interaction of treatment with the risk linear predictor:

$$P(y=1\given lp, t_x;\hat{\bm{\beta}}) = g(\delta_0 + \delta_1\hat{lp} +
\gamma_0t_x + \gamma_1\hat{lp}t_x)$$

Then the predicted absolute benefit is 
$$\tau(lp;\hat{\bm{\beta}},\bm{\gamma}, \bm{\delta}) = g(\delta_0 +
\delta_1\hat{lp}) - g\big((\delta_0+\gamma_0) + (\delta_1+\gamma_1)\hat{lp}\big)$$

## Restricted cubic splines
Finally, we drop the linearity assumption and predict absolute benefit using
smoothing with restricted cubic splines with $k=3, 4$ and $5$ knots. More
specifically, we fit the model:

$$P(y=1|lp, t_x;\hat{\bm{\beta}}, \bm{\delta}, f_{RCS}) = g\big(\delta_0 + \gamma
t_x +f_{RCS}(\hat{lp})\times t_x\big)$$
where
$$f_{RCS}(x)=\alpha_0+\alpha_1h_1(x)+\alpha_2h_2(x)+\dots+\alpha_{k-1}h_{k-1}(x)$$
with $h_1(x)=x$ and for $j=2,\dots,k-2$
$$h_{j+1}(x)= (x-t_j)^3-(x-t_{k-1})_+^3 \frac{t_k-t_j}{t_k-t_{k-1}}+(x-t_k)^3_+\frac{t_{k-1}-t_j}{t_k-t_{k-1}}$$
where 
$t_1,\dots,t_k$ are the selected knots. We predict absolute benefit from 
$$ \tau(lp; \hat{\bm{\beta}}, \delta_0, f_{RCS}) = P(y=1\given\hat{lp},
t_x=0) - P(y=1\given \hat{lp}, t_x=1) $$

\newpage
# Adaptive model selection frequencies

```{r adaptiveBase, cache=TRUE, echo=FALSE, fig.cap="Model selection frequencies of the adaptive approach based on Akaike's Information Criterion across 500 replications. The scenario with the true constant relative treatment effect (first panel) had a true prediction AUC of 0.75 and sample size of 4,250. ", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/selected_model_adaptive_base.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

```{r adaptiveSampleSize, cache=TRUE, echo=FALSE, fig.cap="Model selection frequencies of the adaptive approach based on Akaike's Information Criterion across 500 replications. Sample size is 17,000 rather than 4,250 in Figure \\ref{fig:adaptiveBase}", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/selected_model_adaptive_sample_size.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

```{r adaptiveAuc, cache=TRUE, echo=FALSE, fig.cap="Model selection frequencies of the adaptive approach based on Akaike's Information Criterion across 500 replications. AUC is 0.85 rather than 0.75 in Figure \\ref{fig:adaptiveBase}", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/selected_model_adaptive_auc.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

\newpage
# Discrimination and calibration for benefit

```{r discriminationSampleSize, cache=TRUE, echo=FALSE, fig.cap="Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.75 and sample size of 17,000", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/discrimination_moderate_sample_size.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

```{r discriminationAuc, cache=TRUE, echo=FALSE, fig.cap="Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.85 and sample size of 4,250", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/discrimination_moderate_auc.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

```{r calibrationSampleSize, cache=TRUE, echo=FALSE, fig.cap="Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.75 and sample size of 17,000", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/calibration_moderate_sample_size.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

```{r calibrationAuc, cache=TRUE, echo=FALSE, fig.cap="Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. True prediction AUC of 0.85 and sample size of 4,250", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/calibration_moderate_auc.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

\newpage
# Strong relative treatment effect

Here we present the root mean squared error of the considered methods using
strong constant relative treatment effect ($\text{OR}=0.5$) as the
reference. Again, the same sample size and prediction performance settings were
considered along with the same settings for linear, quadratic and non-monotonic
deviations from the base case scenario of constant relative treatment effects
are considered. All results can be found at
[https://arekkas.shinyapps.io/simulation_viewer/](https://arekkas.shinyapps.io/simulation_viewer/).

```{r rmseHighBase, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated super-population of size 500,000. The scenario with true constant relative treatment effect (panel A) had a true prediction AUC of 0.75 and sample size of 4,250. The RMSE is also presented for strong linear (panel B), strong quadratic (panel C), and non-monotonic (panel D) deviations from constant relative treatment effects. Panels on the right side present the true relationship between baseline risk (x-axis) and absolute treatment benefit (y-axis). The 2.5, 25, 75 and 97.5 percentiles of the risk distribution are expressed in the boxplot.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_high_base.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

```{r rmseHighSampleSize, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replcations calculated in a simulated sample of size 500,000. Sample size is 17,000 rather than 4,250 in Figure \\ref{fig:rmseHighBase}.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_high_sample_size.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

```{r rmseHighAuc, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replcations calculated in a simulated sample of size 500,000. AUC is 0.85 rather than in Figure \\ref{fig:rmseHighBase}.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_high_auc.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

\newpage
# Treatment interactions

We carried out a smaller set of simulations, in which we assumed true
treatment-covariate interactions. Sample size was set to 4,250 and the AUC of
the true prediction model was set to 0.75. The following scenarios were
considered: 1) 4 true weak positive interactions
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.83$); 2) 4 strong positive
interactions ($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.61$); 3) 2 weak and 2
strong positive interactions; 4) 4 weak negative interactions 
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=1.17$); 5) 4 strong negative interactions 
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=1.39$); 6) 2 weak and 2 strong negative
interactions; 7) combined positive and negative strong interactions. We also
considered constant treatment-related harms applied on the absolute scale to all
treated patients. The exact settings were: 1) absent treatment-related harms; 2)
moderate treatment-related harms, defined as 25\% of the average true benefit of the
scenario without treatment-related harms; 3) strong treatment-related harms
defined as 50\% of the true average benefit of the scenario without
treatment-related harms; 4) negative treatment-related harms (benefit), defined
as an absolute risk reduction for treated patients of 50\% of the true average
benefit of the scenario without treatment-related harms. The exact settings can
be found in Table \@ref(tab:tab2).

```{r tab2, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
asd <- read_csv(
  here("data/processed/analysisIdsInteractions.csv")
) %>%
  select(-"type")
kableExtra::kbl(
  asd,
  format = "latex",
  escape = FALSE,
  longtable = TRUE,
  align = "r",
  booktabs = TRUE,
  caption = "Scenario settings of the treatment-covariate interaction scenarios.",
  col.names = c(
    "Scenario",
    "Base",
    "N",
    "AUC",
    "Treatment-related harm",
    paste0("b", 0:8),
    "g0", "g1", "g2", "g5", "g6",
    "Before harms",
    "After harms"
  ),
  linesep = c('', '', '', '\\addlinespace'),
  digits      = 2,
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::landscape() %>%
  kableExtra::add_header_above(
    c(
      "Analysis ID"           = 5,
      "Baseline risk"         = 9,
      "True treatment effect" = 5,
      "Benefit"               = 2
      )
  ) %>%
  kableExtra::kable_styling(font_size = 6)
```

```{r rmseInteractionPositive, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000 where treatment-covariate interactions all favoring treatment were considered.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_interaction_positive.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```
```{r rmseInteractionNegative, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000 where treatment-covariate interactions all favoring the control were considered.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_interaction_negative.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```
```{r rmseInteractionCombined, cache=TRUE, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000 where treatment-covariate interactions 2 favoring treatment and 2 favoring the control were considered.", fig.show="hold", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here("figures/rmse_interaction_combined.tiff")))
# knitr::include_graphics(here::here("figures/calibration_base.png"))
```

\newpage
# Empirical illustration

For predicting baseline risk of 30-day mortality we fitted a logistic regression
model with age, Killip class (*Killip*), systolic blood pressure (*sysbp*),
pulse rate (*pulse*), prior myocardial infarction (*pmi*), location of
myocardial infarction (*miloc*) and treatment as the covariates. Baseline
predictions were made setting treatment to 0.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
load(here::here("data/raw/gusto.rda"))
gusto <- gusto %>%
  tibble() %>%
  filter(!is.na(tpa))

gusto <- gusto %>%
  tibble() %>%
  filter(tx != "SK+tPA") %>%
  rename(
    "outcome" = "day30",
    "treatment" = "tpa"
  )

treatmentArms <- gusto %>%
  group_by(treatment) %>%
  summarise(n = n())

prediction <- glm(
   outcome ~ treatment + age + Killip + pmin(sysbp, 120) + lsp(pulse, 50) + pmi + miloc,
   data = gusto,
   family = "binomial",
   maxit = 99
)

predictionCoefficients <- coef(prediction)
predictionCoefficients <- predictionCoefficients[c(1, 3:12, 2)]
```

\begin{equation*}
P(outcome=1|X=x) = \text{expit}(lp(x)),
(\#eq:gusto1)
\end{equation*}
where 
\begin{equation*}
\begin{aligned}
lp(x)=& \beta_0 + \beta_1 \text{age} + \beta_2 I(\text{Killip}=II) + \beta_3I(\text{Killip}=III) +\\
&\beta_4 I(\text{Killip}=IV) + \beta_5min(\text{sysbp}, 120) + \beta_6 \text{pulse}+\\
&\beta_7 max(\text{pulse - 50, 0}) + \beta_8 I(\text{pmi}=yes)+\\
&\beta_9 I(\text{miloc}=Anterior) + \beta_9 I(\text{miloc}=Other) +\\
&\gamma\times\text{treatment}
\end{aligned}
(\#eq:gusto2)
\end{equation*}

and $expit(x) = \frac{e^x}{1+e^x}$

```{r, echo=FALSE}
predictionSummary <- summary(prediction)
# predictionSummaryCoef <- tibble(predictionSummary$coefficients) %>%
#   rownames_to_column("Variable")

predictionCoefficientsTable <- predictionSummary$coefficients %>%
  as.data.frame() %>%
  mutate(
    rownames = c(
      "Intercept",
      "Age",
      "Killip class = II",
      "Killip class = III",
      "Killip class = IV",
      "Systolic blood pressure",
      "Pulse rate (1)",
      "Pulse rate (2)",
      "Previous MI (yes)",
      "MI location (Other)",
      "MI location (Anterior)",
      "Treatment"
    )
  ) %>%
  rownames_to_column("Variable") %>%
  select(-Variable) %>%
  select(rownames, everything())

colnames(predictionCoefficientsTable) <- c(
  "Variable",
  "Estimate",
  "stderror",
  "zvalue",
  "pvalue"
)

kableExtra::kbl(
  predictionCoefficientsTable,
  format = "latex",
  escape = FALSE,
  longtable = TRUE,
  align = c("l", rep("r", 4)),
  booktabs = TRUE,
  caption = "Coefficients of the prediction model for 30-day mortality, based on the data from GUSTO-I trial.",
  digits = 3,
  linesep = "",
  format.args = list(
    big.mar      = ",",
    decimal.mark = "."
  )
) %>%
  kableExtra::kable_styling(font_size = 7)
```