### 2022-02-16 submission/iscb_2022_abstract.rmd
add outline of iscb abstract

---
title: |
 Individualized treatment effect was predicted best by modeling 
 baseline risk in interaction with treatment assignment
abstract: |
author:
  - name: Alexandros Rekkas
    affiliation: 1
  - name: Peter R. Rijnbeek
    affiliation: 1
  - name: David M. Kent
    affiliation: 2
  - name: Ewout W. Steyerberg
    affiliation: 3
  - name: David van Klaveren
    affiliation: 4
address:
  - code: 1
    address: Department of Medical Informatics, Erasmus Medical Center, Rotterdam, The Netherlands
  - code: 2
    address: Predictive Analytics and Comparative Effectiveness Center, Institute for Clinical Research and Health Policy Studies, Tufts Medical Center, Boston, Massachusetts, USA
  - code: 3
    address: Department of Biomedical Data Sciences, Leiden University Medical Center, Leiden, The Netherlands
  - code: 4
    address: Department of Public Health, Erasmus Medical Center, Rotterdam, The Netherlands
numbersections: true
linenumbers: false
journal: arXiv.org
keywords:
  - treatment effect heterogeneity
  - absolute benefit
  - prediction models
output: word_document
geometry: margin=1.0in
date: false
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \usepackage{amsmath}
  - \doublespacing
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \date{}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(SmoothHte)
library(rms)
library(here)

d <- function(x, decimals = 4) {
  sprintf(paste0("%1.", decimals, "f"), x) 
}

knit_hooks$set(
  inline = function(x) {
    prettyNum(
      x,
      big.mark = ",",
      decimal.mark = ".",
      preserve.width = "individual"
    )
  }
  )
set.seed(19910930)
```

# Abstract
Our objective was to compare different risk-based methods for optimal prediction
of individualized treatment effects from RCTs.

We simulated RCT data using diverse assumptions for the average treatment
effect, a baseline prognostic index of risk (PI), the shape of its interaction
with treatment (none, linear, quadratic or non-monotonic) and the magnitude of
treatment-related harms (none or constant independent of the PI). The
combination of these settings resulted in the definition of 648 simulation
scenarios. In each sample we predicted absolute benefit using: models with a
constant relative treatment effect; stratification in quarters of the PI; models
including a linear interaction of treatment with the PI; models including an
interaction of treatment with a restricted cubic spline (RCS) transformation of
the PI; an adaptive approach using Akaike’s Information Criterion. We evaluated
predictive performance using root mean squared error and measures of
discrimination and calibration for benefit.

The linear-interaction model and the RCS-interaction model outperformed the
constant treatment effect model in many simulation scenarios. The RCS-model was
optimal when quadratic or non-monotonic deviations from a constant treatment
effect were stronger, and when sample size was larger. Larger sample size also
supported the adaptive approach. All the simulation results can be explored at
https://arekkas.shinyapps.io/simulation_viewer/. We also illustrated the
application of the considered methods in the GUSTO-I trial, which confirmed our
simulations' findings.

In conclusion, an interaction between baseline risk and treatment assignment
should be considered to improve treatment effect predictions. Non-linear
interactions should be considered only in larger sample sizes.